A large leading tail of atmospheric escape in the HAT-P-67 b system

### 2023-05-15 paper/paper1/ms.tex
Will send a draft to Caroline today, reading through now.

\keywords{stars: fundamental parameters ---  stars: statistics}

@@

A runaway inflation scenario predicts an inevitable and profound mass loss rate for inflated hot Saturns.  As the planet's atmosphere overflows its Roche lobe at an ever-increasing pace instantaneous mass loss rates may exceed $\dot{M}>10^{13}$, about 10$\times$ larger than those previously seen in atmospheric escape measurements to date \citep{2022arXiv221116243D}.  Inflated hot Saturns therefore make excellent targets for direct measurement of atmospheric escape.

Large mass loss alone does not guarantee detectability.  The ability to detect even immense mass loss hinges on its observability in spectral tracers.  Ly$\alpha$, \ion{He}{1} 10833 \AA, and H$\alpha$ have emerged as the most amenable to detection \citep{2018ApJ...855L..11O,2022arXiv221116243D,2023MNRAS.518.4357O}, but each of these has its own observational limitations.  Ly $\alpha$ can suffer from interstellar medium \ion{H}{1} absorption censoring its low-velocity line core, for example.  Here we focus on \ion{He}{1} 10833 \AA, which offers some advantages.  In particular, metastable Helium's ability to be observed at high spectral resolution from the ground has been especially valuable for evincing velocity substructure of the escaping gas motion relative to the exoplanet restframe \citep{2019A&A...629A.110A,2020ApJ...894...97N}, and convincingly associating the signal to an exoplanetary origin as opposed to stellar contamination \citep{2018AJ....156..189C}.  \ion{He}{1} has resulted in at least 14 systems with detections, including \object[GJ 3470]{GJ3470} \citep{2020ApJ...894...97N, 2021A&A...647A.129L}, \object{HD 63433 c} \citep{2022AJ....163...68Z}, WASP 107 b \citep{2019A&A...623A..58A,2020AJ....159..115K}, HD 209458 b \citep{2019A&A...629A.110A}, WASP 69 b \citep{2020AJ....159..278V}, WASP 52 b \citep{2020AJ....159..278V}, HD 189733 b \citep{2021A&A...647A.129L}, and HAT-P-18 b \citep{2021ApJ...909L..10P}.  An observational picture of mass loss in hot Saturns appears to be lacking, at least in part due to their intrinsic rarity.

@@

Here we present a multi-year observational campaign searching for atmospheric escape in \object{HAT-P-67 b}, an inflated hot Saturn orbiting an F type sub-giant at an orbital separation of 0.06 AU and a 4.8-day orbital period \citep{2017AJ....153..211Z}.  Its strong insolation ($T_\mathrm{eq}\sim2000\;$K) combined with \object{HAT-P-67 b}'s extremely low surface gravity makes it an exceptional candidate for strong atmospheric mass loss via Roche lobe overflow.  Its $\sim$0.05 g~cm$^{-3}$ density places it below the 1.0 g~cm$^{-3}$ threshold predicted to exhibit runaway inflation by \citet{2023ApJ...945L..36T}.  Figures \ref{fig:instellation} and \ref{fig:ThornFortUpdated} show how much of an outlier \hatp is: large, low mass, and heavily irradiated.

@@

Only 19 out of the 39 nights possess A0V telluric calibration standard stars.  These standard stars were used to spot-check our telluric masking.  Figure \ref{TESSoverview} indicates the epochs of select HPF visits overlaid on the \emph{TESS} time-series lightcurve.

@@

The 13.8 hours of exposure, combined with \object{HAT-P-67 b}'s short ($P\sim4.8$ day) period, means that a large fraction of the orbit has been collected, with some phases (especially near transit) heavily sampled, and some other out-of-transit phases sampled more sparsely. For visualization purposes, we constructed a 2-D intensity phase scan, binning in phase and wavelength, and spanning the entire orbit of \object{HAT-P-67 b}.  Figure \ref{fig:HPFscanResid} overlays the planet's approximately $\pm150\;$km/s orbital Doppler velocity, with the stellar rest frame velocity demarcated by the vertical line.  The vanishing $<36$ m/s reflex motion of the star is imperceptible at this scale.  The horizontal dashed lines indicate the moments of transit ingress (-0.03), mid-center (0.0), and egress (+0.03).

@@

The predicted Helium ionization depends sensitively on the XUV flux \citep{2019ApJ...881..133O}.  Ideally, we would have a measurement of the XUV spectrum of \hatp, but the limited facilities and distance of the source preclude these challenging observations.  Instead, we constructed panchromatic X-ray to visible SEDs by scaling and stitching together synthetic spectra to provide a range of high and low $L_X/L_\mathrm{bol}$.  We chose a 6400 K PHOENIX \citep{husser13} solar metallicity photospheric spectral template with $\log{g}=3.5$, scaled to the solid angle seen by \hatpb.  The F7IV-V \object[tau Boo]{$\tau$ Boo} serves as the closest analog with available synthetic x-ray coronal spectra \citep{2011A&A...532A...6S}.  But \hatp resides closer to the Kraft break than does \object[tau Boo]{$\tau$ Boo}, and the move to hotter F stars may be associated with weakening of the stellar wind and other atmospheric changes leading to the prospect of much lower XUV luminosity for HAT-P-67 \citep{2022ApJ...930....7A}.  So there remains significant uncertainty about the applicability of the \object[tau Boo]{$\tau$ Boo} XUV spectrum to HAT-P-67.  We scaled the synthetic SED of \object[tau Boo]{$\tau$ Boo} retrieved from X-exoplanets \citep{2011A&A...532A...6S} so that the integral of ground-state ionizing photons ($\lambda<504\AA$) exhibited $L_X/L_\mathrm{bol} \in [10^{-7}, 10^{-5}]$.  Figure \ref{fig:XUV} shows two conceivable SEDs with these high and low radiation hardness.

@@

@@

Second, the star's non-negligible rotational velocity sets up a configuration analogous to Doppler Tomography in the Rossiter-McLaughlin (RM) effect, but distinct in a subtle way.  Whereas traditional Doppler Tomography treats the scanning reticle as an opaque planetary disk, here the reticle resembles a transmissive filter with a wavelength center and width that varies in space and time.  We rank this effect as second-order and ignore it for now.
%%We can understand this transmissive Doppler tomography effect through an example of a pathological limiting case.  Imagine the escaping atmsosphere were concentrated in a 50\% transmissive cylinder coming out of the orbital plane, as in an idealized transparent bipolar outflow.  This cyclinder would possess the planet's bulk velocity of about $-30$ km/s at the moment of transit ingress, placing the locus of absorption at about 10832.4 \AA.  But the absorptive cylinder would intercept a finite range of stellar limb velocities, broadening the apparent width of the absorption line.

@@

Assuming $\log{L_\mathrm{XUV}/L_\mathrm{bol}}=-4.2$, we would expect \hatpb~ to exhibit $\dot{M}\sim10^2$ M$_\oplus$/Gyr. At a current $M\sim95\;$M$_\oplus$ and an accelerating mass loss rate, its lifetime would measure in the $\sim$100 Myr range.  The \citet{2023ApJ...945L..36T} simulations focused on 0.75-1.25 $M_\odot$ host stars; the higher mass $\sim$1.6 $M_\odot$ \hatp would deliver greater tidal gravity for a given semi-major axis, and possibly lower $\log{L_\mathrm{XUV}/L_\mathrm{bol}}$ compared to more active G stars.  Nevertheless, we can project the time series trends in their Figures 3 and 4 to recreate a qualitative history and fate of \hatpb~ under the assumptions of XUV photoevaporation.  Broadly, the planet would have started with an initial mass up to 50$\%$ more massive than it is now, with $R\sim\;$1.4 $R_\mathrm{Jup}$, for an initial density of $\sim$0.2 g/cm$^-3$.  It would lose mass at a rate of tens of Earth masses per Gyr for its 1 Gyr lifetime, expanding modestly until it reaches the critical $\sim$0.1 g/cm$^-3$ threshold, at which point the mass loss rate increases to its current value.  Having lost a significant portion of its mass, its orbit expands to conserve angular momentum.  It will last only tens of Myr in its current state before losing almost all of its envelope and settling as a $5-15$ M$_\oplus$ core with a final radius of 0.2$-$0.3 R$_\mathrm{Jup}$.

@@

\citet{2017AJ....153..211Z} estimated that \hatpb~received about twice the incident flux as a Zero Age Main Sequence (ZAMS) \hatp, based on comparison to the Geneva isochrones. In Section \ref{secMISTtracks} MIST evolutionary tracks showed two equally plausible states: the tail-end of the main sequence or a recently evolved sub-giant.  Figure X employs these tracks to assess what the impact on \hatpb~could be, assuming that the orbital location has not changed over the system lifetime (\emph{i.e.} star-planet or planet-planet migration has been negligible).  The second panel from the top illustrates a countervailing effect: the decrease in Ohmic dissipation heating efficiency, $\epsilon$, could counteract the increase of stellar flux, assuming the functional form for $\hat{\epsilon}_G$ described by \citet{2018AJ....155..214T}.  Together the effects nearly cancel when computing the inflationary lifetimes, yielding nearly identical curves in the bottom panel.  The third panel recreates the numerically computed curve for $0.5\;M_\mathrm{Jup}$, $T_\mathrm{eq}=$1800 K from \citep{2011ApJ...738....1B}, which should be considered as representative since it was not necessarily tailored to the evolutionary history of \hatpb.  Nevertheless, the similarity of the curve is remarkable since it arrives at approximately the correct planet radius at the right age.  The modeled mass is slightly higher than, but consistent with the RV-measured mass, and could imply a large and increasing mass loss rate over its lifetime.

@@

We, therefore, attribute this cliff in the $M_p$ vs $R_p$ diagram to Ohmic dissipation.  \citet{2018AJ....155..214T} emphasized the positive feedback loop as a cause for the lack of sub-Saturns and noted an alternative interpretation-- mass-dependent migration segregation.  While both scenarios are conceivable, the evidence of strong evaporation of \hatpb~ favors the evaporation scenario.  Furthermore, the runaway evaporation of hot Saturns was naturally predicted by the Ohmic dissipation mechanism, whereas mass-dependent migration has only been developed as a post-diction.

@@

\section{Discussion}
\subsection{Orbital constraints}
We premised our joint \emph{TESS} lightcurve and HPF RV time series orbit fit on a circular orbit.

\subsection{Cause for stellar modulation}
As discussed in Section \ref{TESSmodulation}, the \emph{TESS} lightcurve exhibits up to $0.36\%$ peak-to-valley modulation amplitude, with a characteristic timescale of $P=4.7-5.9$ days. Some sectors exhibit lower amplitudes.  The conventional interpretation would consign these cyclical modulations to the familiar stellar activity: surface features---either starspots, faculae, or plage---entering and exiting the projected stellar disk on the stellar rotation period comparable to the 4.8-day orbital period of planet b.  The cause for such orbital and rotational synchronization would then ostensibly be star-planet tidal interactions.  We examine the prospect of tides in the next subsection.

Alternatively, the extreme mass loss and credibility of the Ohmic dissipation model motivate the prospect of Star Planet Magnetic Interaction (SPMI) that we consider now.  In this scenario, the magnetic field of the planet permeates the space between the star and the planet.  Magnetic perturbations propagate via planetary magnetic fields with the Alfv\'en speed.  The planet can interact with the star if the Alfv\'en speed exceeds the stellar wind speed controlling the bulk motion of the intervening medium.  The magnetic field of the planet is not known, but the ohmic dissipation mechanism favors a limited range in the magnetic field strength: large enough to cause significant dissipative energy generation, but not too large to shut down charged particle motion entirely.  \citet{2011ApJ...738....1B} estimate this range as between about 1-30~G.

The non-detection of variability in the \ion{Ca}{2} H and K lines and H~$\alpha$ lines appears to disfavor this SPMI interpretation.

\subsection{Extended Helium exospheres in non-transiting planets}

We measured a large extent of Helium escape along the arc of the orbit, but we necessarily can place only coarse constraints on the vertical extent--in the direction perpendicular to the orbital path, $\pm z$.  We estimate the detectable vertical extent must be at least a few stellar radii, such that we still could have detected Helium escape in a hypothetical HAT-P-67b-like system even if it were non-transiting, in a ``near miss'' configuration.  We propose a new category of semi-transiting planet, dubbed ``exospheric grazers'', in which the planet does not produce a detectable white-light transit depth, but does produce measureable \emph{line}-based exospheric absorption.  This category appears to have been neglected due to the assumption that exospheres were only easily detectable at several planetary radii, and therefore fractions of a stellar radius.  While large tails have been seen previously in Ly$\alpha$, the scarcity of UV resources prohibited searches of this kind.  The discovery of large Helium tails in HAT-P-32b (Zhang et al. 2023) and HAT-P-67b suggest that the identification of these exospheric grazers may be achievable with current instrumentation and possibly existing archival data from near-IR RV-planet searches.

@@

A few benefits of of this planet category arise.  First, they would expand the sample of planets for which the planetary winds could be measured, by relaxing the narrow geometrical transit probability of $1-10^\circ$ to something closer to $\pm30^\circ$, depending on the mass loss and gas dynamics.  In particular, the smallest stars such as white dwarfs and M-dwarfs could see the search space for exospheres increase by $10-100\times$.  Second, the range of viewing angles would serve as a natural experiment for constraining the vertical extent of outflows in the first place.  The typical bow shock 3D morphology could be deduced from a sample of otherwise-similar planets with a range of orbital inclinations.  Non-detections would therefore be valuable, especially in cases where $\sin{i}$ could be constrained independently.  In practice these additional constraints are rarely possible, so ensemble statistical methods may have to be employed.  Finally, for some promising planet candidates a distinctive helium signal could validate the planet, assuming the signal could be disambiguated from stellar variability.  That disambiguation may be feasible, since the velocity structure and phasing of stellar and planetary Helium signals diverge due to stellar wind confinement and a launching from a predictable planetary rest frame.  The elusive category of young planets could be ideal, since their large radii and high intrinsic temperatures and high incident XUV radiation may yield large mass loss rates if they have recently migrated to an interior location through disk migration.  This category is especially valuable if discovered through near-IR RV searches, since abundant multi-epoch spectroscopy may already contain the requisite stellar variability baselines.

>>>>

\keywords{Exoplanet atmospheres, Exoplanet evolution, Exoplanet atmospheric dynamics, Stellar winds, Exoplanet atmospheric variability}

@@

A runaway inflation scenario predicts an inevitable and profound mass loss rate for inflated hot Saturns.  As the planet's atmosphere overflows its Roche lobe at an ever-increasing pace instantaneous mass loss rates may exceed $\dot{M}>10^{13}$ g/s, over 10$\times$ larger than those previously seen in atmospheric escape measurements to date \citep{2022arXiv221116243D}.  Inflated hot Saturns therefore make excellent targets for direct measurement of atmospheric escape.

Large mass loss alone does not guarantee detectability.  The ability to detect even immense mass loss hinges on its observability in spectral tracers.  Ly$\alpha$, \ion{He}{1} 10833 \AA, and H$\alpha$ have emerged as the most amenable to detection \citep{2018ApJ...855L..11O,2022arXiv221116243D,2023MNRAS.518.4357O}, but each of these has its own observational limitations.  Ly$\alpha$ can suffer from interstellar medium \ion{H}{1} absorption censoring its low-velocity line core, for example.  Here we focus on \ion{He}{1} 10833 \AA, which offers some advantages.  In particular, metastable Helium's ability to be observed at high spectral resolution from the ground has been especially valuable for evincing velocity substructure of the escaping gas motion relative to the exoplanet restframe \citep{2019A&A...629A.110A,2020ApJ...894...97N}, and convincingly associating the signal to an exoplanetary origin as opposed to stellar contamination \citep{2018AJ....156..189C}.  \ion{He}{1} has resulted in at least 14 systems with detections, including \object[GJ 3470]{GJ3470} \citep{2020ApJ...894...97N, 2021A&A...647A.129L}, \object{HD 63433 c} \citep{2022AJ....163...68Z}, WASP 107 b \citep{2019A&A...623A..58A,2020AJ....159..115K}, HD 209458 b \citep{2019A&A...629A.110A}, WASP 69 b \citep{2020AJ....159..278V}, WASP 52 b \citep{2020AJ....159..278V}, HD 189733 b \citep{2021A&A...647A.129L}, and HAT-P-18 b \citep{2021ApJ...909L..10P}.  An observational picture of mass loss in hot Saturns appears to be lacking, at least in part due to their intrinsic rarity.

@@

Here we present a multi-year observational campaign searching for atmospheric escape in \object{HAT-P-67 b}, an inflated hot Saturn orbiting an F type sub-giant at an orbital separation of 0.06 AU and a 4.8-day orbital period \citep{2017AJ....153..211Z}.  Its strong insolation ($T_\mathrm{eq}\sim2000\;$K) combined with \object{HAT-P-67 b}'s extremely low surface gravity makes it an exceptional candidate for strong atmospheric mass loss via Roche lobe overflow.  Its $\sim$0.05 g~cm$^{-3}$ density places it firmly below the 1.0 g~cm$^{-3}$ threshold predicted to exhibit runaway inflation by \citet{2023ApJ...945L..36T}.  Figures \ref{fig:instellation} and \ref{fig:ThornFortUpdated} show how much of an outlier \hatpb~ is: large, low mass, and heavily irradiated.

@@

Only 19 out of the 39 nights possess A0V telluric calibration standard stars.  These standard stars were used to spot-check our telluric masking.  Figure \ref{fig:TESSoverview} indicates the epochs of select HPF visits overlaid on the \emph{TESS} time-series lightcurve.

@@

The 13.8 hours of exposure, combined with \object{HAT-P-67 b}'s short ($P\sim4.8$ day) period, means that a large fraction of the orbit has been collected, with some phases (especially near transit) heavily sampled, and some other out-of-transit phases sampled more sparsely. For visualization purposes, we constructed a 2-D intensity phase scan, binning in phase and wavelength, and spanning the entire orbit of \object{HAT-P-67 b}.  Figure \ref{fig:HPFscanResid} overlays the planet's approximately $\pm150\;$km/s orbital Doppler velocity, with the stellar rest frame velocity demarcated by the vertical line.  The vanishing $<36$ m/s reflex motion of the star is imperceptible at this scale.  The horizontal dashed lines indicate the moments of transit ingress (-0.03), mid-center (0.0), and egress (+0.03), demarcated as \texttt{TRANSIT} in the figure.

We define 4 additional distinct groups of phases with adjectival qualifiers in anticipation of the need to discuss bulk trends: \texttt{PRE}, \texttt{POST}, \texttt{CONTROL}, and \texttt{BASELINE}.  The \texttt{BASELINE} phases define the out-of-transit baseline.  The ``control'' spectra designate phases just before the baseline, but not used to define the baseline, to inspect minor correlation structure in the spectra without dividing it out.  Table \ref{tabHPFLog} lists the adjectival qualifiers for each HPF visit.

@@

The predicted Helium ionization depends sensitively on the XUV flux \citep{2019ApJ...881..133O}.  Ideally, we would have a measurement of the XUV spectrum of \hatp, but the limited facilities and distance of the source preclude these challenging observations.  Instead, we constructed panchromatic X-ray to visible SEDs by scaling and stitching together synthetic spectra to provide a range of high and low $L_X/L_\mathrm{bol}$.  We chose a 6400 K PHOENIX \citep{husser13} solar metallicity photospheric spectral template with $\log{g}=3.5$, scaled to the solid angle seen by \hatpb.  The F7IV-V \object[tau Boo]{$\tau$ Boo} serves as the closest analog with available synthetic X-ray coronal spectra \citep{2011A&A...532A...6S}.  But \hatp resides closer to the Kraft break than does \object[tau Boo]{$\tau$ Boo}, and the move to hotter F stars may be associated with weakening of the stellar wind and other atmospheric changes leading to the prospect of much lower XUV luminosity for HAT-P-67 \citep{2022ApJ...930....7A}.  So there remains significant uncertainty about the applicability of the \object[tau Boo]{$\tau$ Boo} XUV spectrum to HAT-P-67.  We scaled the synthetic SED of \object[tau Boo]{$\tau$ Boo} retrieved from X-exoplanets \citep{2011A&A...532A...6S} so that the integral of ground-state ionizing photons ($\lambda<504\AA$) exhibited $L_X/L_\mathrm{bol} \in [10^{-7}, 10^{-5}]$.  Figure \ref{fig:XUV} shows two conceivable SEDs with these high and low radiation hardness.

@@

\begin{figure}
    \includegraphics[width=0.8\linewidth]{figures/homogeneous.png}
    \caption{Isotropic mass loss}
    \label{fig:IsotropicMdot}
\end{figure}

\begin{figure}
    \includegraphics[width=0.8\linewidth]{figures/asym.png}
    \caption{Asymmetric dayside mass loss}
    \label{fig:DaysideMdot}
\end{figure}

@@

Second, the star's non-negligible rotational velocity sets up a configuration analogous to Doppler Tomography in the Rossiter-McLaughlin (RM) effect, but distinct in a subtle way.  Whereas traditional Doppler Tomography treats the scanning reticle as an opaque planetary disk, here the reticle resembles a transmissive filter with a wavelength center and width that varies in space and time.  We rank this effect as second-order and return to it in the Discussion Section \ref{vsini-interplay}.

@@

Assuming $\log{L_\mathrm{XUV}/L_\mathrm{bol}}=-4.2$, we would expect \hatpb~ to exhibit $\dot{M}\sim10^2$ M$_\oplus$/Gyr. At a current $M\sim95\;$M$_\oplus$ and an accelerating mass loss rate, its lifetime would measure in the $\sim$100 Myr range.  The \citet{2023ApJ...945L..36T} simulations focused on 0.75-1.25 $M_\odot$ host stars; the higher mass $\sim$1.6 $M_\odot$ \hatp would deliver greater tidal gravity for a given semi-major axis, and possibly lower $\log{L_\mathrm{XUV}/L_\mathrm{bol}}$ compared to more active G stars.  Nevertheless, we can project the time series trends in their Figures 3 and 4 to recreate a qualitative history and fate of \hatpb~ under the assumptions of XUV photoevaporation.  Broadly, the planet would have started with an initial mass up to 50$\%$ greater than at present, with $R\sim\;$1.4 $R_\mathrm{Jup}$, for an initial density of $\sim$0.2 g/cm$^-3$.  It would lose mass at a rate of tens of Earth masses per Gyr for its 1 Gyr lifetime, expanding modestly until it reaches the critical $\sim$0.1 g/cm$^-3$ threshold, at which point the mass loss rate increases to its current value.  Having lost a significant portion of its mass, its orbit expands to conserve angular momentum.  It will last only tens of Myr in its current state before losing almost all of its envelope and settling as a $5-15$ M$_\oplus$ core with a final radius of 0.2$-$0.3 R$_\mathrm{Jup}$.

@@

\citet{2017AJ....153..211Z} estimated that \hatpb~received about twice the incident flux as a Zero Age Main Sequence (ZAMS) \hatp, based on comparison to the Geneva isochrones. In Section \ref{secMISTtracks} MIST evolutionary tracks showed two equally plausible states: the tail-end of the main sequence or a recently evolved sub-giant.  Figure \ref{fig:OhmicInflate} employs these tracks to assess the re-inflationary impact of \hatpb~could be, assuming that the orbital location has not changed over the system lifetime.  The second panel from the top illustrates a countervailing effect: the decrease in Ohmic dissipation heating efficiency, $\epsilon$, could counteract the increase of stellar flux, assuming the functional form for $\hat{\epsilon}_G$ described by \citet{2018AJ....155..214T}.  Together the effects nearly cancel when computing the inflationary lifetimes, yielding nearly identical curves in the bottom panel.  The third panel recreates the numerically computed curve for $0.5\;M_\mathrm{Jup}$, $T_\mathrm{eq}=$1800 K from \citep{2011ApJ...738....1B}, which should be considered as representative since it was not necessarily tailored to the evolutionary history of \hatpb.  Nevertheless, the similarity of the curve is remarkable since it arrives at approximately the correct planet radius at the right age.  The modeled mass is slightly higher than, but consistent with the RV-measured mass, and would imply a large and increasing mass loss rate over its lifetime.

@@

We, therefore, attribute this cliff in the $M_p$ vs $R_p$ diagram to either Ohmic dissipation, XUV photoevaporation, or both.  \citet{2018AJ....155..214T} emphasized the positive feedback loop as a cause for the lack of sub-Saturns and noted an alternative interpretation-- mass-dependent migration segregation.  While both scenarios are conceivable, the evidence of strong evaporation of \hatpb~ favors the evaporation scenario.  Furthermore, the runaway evaporation of hot Saturns was naturally predicted by the Ohmic dissipation mechanism, whereas mass-dependent migration has only been developed as a post-diction.

One key observation may tip the scales in favor of XUV photoevaporation as the predominant cause of mass-loss.  The conspicuous leading tail favors preferential mass loss on the highly irradiated planet dayside.  While Ohmic dissipation would exhibit preferential dayside heating, it is the ability of XUV photons to create a planetary wind that tilts the balance.

Finally we consider the prospect that the planetary mass is much lower than its 1-$\sigma$ bound, such that the white-light planetary radius \emph{is} the Hill radius (we have previously assumed the Hill radius is 2.7 planetary radii).  This extreme scenario could manifest enormous mass loss rates, with the planet atmosphere's reservoir of mass fleeing the gravitational potential well directly, without the cushion of an exosphere.  This terminal Roche Lobe overflow would have several observational predictions.  Heavy elements woud easily leak into the planetary wind, yielding possibly many observable metal lines in the UV.  These may be observable even under more mild conditions, due to lofting of metals into the upper atmosphere, so some care would be required to quantitatively distinguish these two scenarios.

@@

\section{Discussion}
%\subsection{Prospects for a non-circular orbit}
%\authorcomment1{Slated for removal:}  
%\textcolor{lightgray}{We premised our joint \emph{TESS} lightcurve and HPF RV time series orbit fit on a circular orbit. Here we consider how a finite eccentricity orbit could effect the observations and lead to the illusion of a leading tail... [...]}

\subsection{Cause for stellar modulation}
As discussed in Section \ref{TESSmodulation}, the \emph{TESS} lightcurve exhibits up to $0.36\%$ peak-to-valley modulation amplitude, with a characteristic timescale of $P=4.7-5.9$ days. Some sectors exhibit lower amplitudes.  The conventional interpretation would consign these cyclical modulations to the familiar stellar activity: surface features---either starspots, faculae, or plage---entering and exiting the projected stellar disk on the stellar rotation period comparable to the 4.8-day orbital period of planet b.  The cause for such orbital and rotational synchronization would then be either coincidence or secular star-planet tidal interactions.  For the latter, disk migration could have naturally ceased near the co-rotation radius, leaving the planet to reside naturally near the stellar $P_\mathrm{rot}$.

Alternatively, Star Planet Magnetic Interaction (SPMI) could be at play.  In this scenario, the magnetic field of the planet permeates the space between the star and the planet.  Magnetic perturbations propagate via planetary magnetic fields with the Alfv\'en speed.  The planet can interact with the star if the Alfv\'en speed exceeds the stellar wind speed controlling the bulk motion of the intervening medium.  The non-detection of variability in the \ion{Ca}{2} H and K lines and H$\alpha$ lines appears to disfavor this SPMI interpretation, since SPMI could be expected to cause variations in these diagnostics.

It is hypothetically possible---albeit unlikely---that mass loss could be great enough to produce variability in the wide band \emph{TESS} lightcurve.  The HPF spectra reveal up to 10\% signal depth over a few Angstroms.  The TESS bandpass barely includes 10830 \AA, at a location of diminishing throughput, so it appears impractical that Helium alone could cause the TESS signal.  Hypothetically, dust dedged up in the mass loss process could cause a large enough broadband flux loss to manifest in TESS.  We may expect to see some variable reddening in that case.  A putative H$\alpha$ line would have to be about 30\% deep over 10 \AA~ wide to manifest perceptibly in the TESS data.  Such an implausibly deep line likely would have been observed as perturbations in the Keck HIRES spectra, even with its limited phasing.

\subsection{Implications for exosphere detection in non-transiting planets}

We measured a large extent of Helium escape along the arc of the orbit, but we necessarily can place only coarse constraints on the vertical extent---in the direction perpendicular to the orbital path, $\pm z$.  We estimate the detectable vertical extent must be at least a few stellar radii, such that we still could have detected Helium escape in a hypothetical HAT-P-67b-like system even if it were non-transiting, in a ``near miss'' configuration.  We propose a new category of semi-transiting planet, dubbed ``exospheric grazers'', in which the planet does not produce a detectable white-light transit depth, but \emph{does} produce measureable \emph{line}-based exospheric absorption.  This category appears to have been neglected due to the assumption that exospheres were only easily detectable at several planetary radii, and therefore fractions of a stellar radius.  While large tails have been seen previously in Ly$\alpha$, the scarcity of UV resources prohibited searches of this kind.  The discovery of large Helium tails in HAT-P-32b (Zhang et al. 2023) and now HAT-P-67b suggest that the identification of these exospheric grazers may be achievable with current instrumentation and possibly existing archival data from near-IR RV-planet searches.

@@

A few benefits of of this planet category arise.  First, they would expand the sample of planets for which the planetary winds could be measured, by relaxing the narrow geometrical transit probability of $1-10^\circ$ to something closer to $\pm30^\circ$, depending on the mass loss and gas dynamics.  In particular, the smallest stars such as white dwarfs and M-dwarfs could see the search space for exospheres increase by $10-100\times$.  Second, the range of viewing angles would serve as a natural experiment for constraining the vertical extent of outflows in the first place.  The typical bow shock 3D morphology could be deduced from a sample of otherwise-similar planets with a range of orbital inclinations.  Non-detections would therefore be valuable, especially in cases where $\sin{i}$ could be constrained independently.  In practice these additional constraints are rarely possible, so ensemble statistical methods may have to be employed.  Finally, for some promising planet candidates a distinctive helium signal could validate the planet, assuming the signal could be disambiguated from stellar variability.  That disambiguation may be feasible, since the velocity structure and phasing of stellar and planetary Helium signals diverge due to stellar wind confinement and the launching from a predictable planetary rest frame.  The elusive category of young planets could make ideal targets, since their large radii and high intrinsic temperatures and high incident XUV radiation may yield large mass loss rates if they have recently migrated to an interior location through disk migration.  This category is especially valuable if discovered through near-IR RV searches, since abundant multi-epoch spectroscopy may already contain the requisite stellar variability baselines.

\subsection{Interplay with rotating stellar disk} \label{vsini-interplay}

As aluded to previously, the interplay between the rotating stellar disk and the Helium line absorption sets up a subtle Doppler tomography-like signal.

We can understand this transmissive Doppler tomography effect through an example of four pathological limiting cases, illustrated in Figure \ref{fig:interplayvsini}.  Imagine the escaping atmsosphere were concentrated in a 20\% transmissive cylinder coming out of the orbital plane, as in an idealized transparent bipolar outflow.  This cyclinder would possess the planet's bulk velocity of about $-30$ km/s at the moment of transit ingress, placing the locus of absorption at about 10832.4 \AA.  But the absorptive cylinder would intercept a finite range of stellar limb velocities.  There would exist a constructive interference at the moment when the planet radial velocity shifts the \ion{He}{1} 10830 \AA line to share the same observed wavelength as the stellar lines at that limb position.

\begin{figure*}[t]
    \includegraphics[width=\linewidth]{figures/disperse_vsini_interplay_v0p1.pdf}
    \caption{Illustration of the interplay between a transiting columnar exosphere (extreme case) and a rotating stellar disk.  The top row shows a non-rotating (negligble $v\sin{i}$) case, with sharp stellar lines.  The motion of the planet can be seen easily, and the signals align at the planet's zero velocity (center of the disk).  The bottom row shows the rapidly rotating case.  Depending on the planet orbital speed and stellar rotational speeds, the signals can align at some pair wavelength before and after transit mid-center.  The alignment of the signals acts to \emph{suppress} the signal since the planet exosphere is blocking the fainter-on-average flux.  If the gas is spread out uniformly over the disk---instead of in a vertical column---the signals simply multiply.}
    \label{fig:interplayvsini}
\end{figure*}
